### 2016-09-08 analysis-note-v2/RTPC.tex
Cleaned up some of the weird and random line jumps

~~~~At 6 GeV incident electron beam energy, the recoil $^{4}He$ nuclei from the coherent DVCS channel, have an average momentum around 200 MeV/c. CLAS cannot detect such low energy particles, for exemple protons are detected with a threshold of about 250 MeV/c, which will be even higher for helium. In order to detect such low-energy recoils a Radial TPC (RTPC), similar to the one used during the BoNuS run \cite{BONUS}, was added to CLAS. In 2009, the CLAS-EG6 group \cite{proposal} has built this RTPC, figure \ref{fig:RTPC2} shows a close photo of the RTPC, while figure \ref{fig:RTPC_xscetion} shows a schematic drawing of its internal structure. It is a 200 mm-long and 150 mm-diameter TPC, composed of two electronically separated modules, each covering about 150$^{\circ}$ of the azimuthal angle. The RTPC tracks the charged particles through the ionization centers they leave in the gas-filled detection volume. Then, the ionized electrons drift under the effect of an electric field and are amplified by a gas electron multiplier system before to be detected on collection pads.

@@

The Ne-DME gas mixture has been chosen as the drift gas because of its 
low-diffusion characteristics and small Lorentz angles (the angles between the 
drift direction of electrons under the influence of magnetic field and the 
direction of the electric field). These characteristics minimize the changes in 
the drift velocity of the ionization electrons \cite{DME}.\\

The GEMs amplify the ionized electrons to produce measurable signals. Figure 
\ref{fig:GEMs} shows a microscopic photo of a GEM. It is made from an insulator 
(Kapton) sandwiched between two copper layers. The mesh of each GEM layer is 
chemically etched with 50-$\mu$m-diameter holes in double-conical cross section 
shapes, as can be seen from the schematic plot on the right.\\

@@

The electrons amplification is achieved through the holes via the strong 
electric field, that is generated by a 400-V potential difference between the 
two copper layers. Such a strong field leads to high ionization of the initial 
electrons and therefore amplification of the signals. Furthermore, an 
additional potential difference of 150 V is set between each two successive GEM 
layers to push the amplified electrons towards the readout board. In this 
configuration, the gain of each GEM layer is of the order of 100.\\

@@

\subsection{Working principle}
~~~~When a charged particle traverses a gas, it ionizes the gas-atoms along its 
trajectory. In a TPC, the electrons released in the ionization drift towards 
the readout board under the effect of an applied electric field. The drift 
velocity depends on the gas mixture, and on the electric ($\vec{E}$) and 
magnetic ($\vec{B}$) fields. The recorded time of the electrons provides 
information on how far the initial ionizations happened in the drift region, 
leading to reconstruct the original points of ionization, while the recorded 
ADCs give the deposited energy.\\

In our TPC the cathode and the anode are two cylinders. Thus, the generated 
$\vec{E}$ field in the drift region has a purely radial components, 
perpendicular to the beam line, with a magnitude around 500 V/cm. However, the 
two-cylinders configuration produces small gradient components at the sides of 
the RTPC. This issue was solved by installing field cages at the ends of the 
RTPC to keep a regular $\vec{E}$ field. These caused some high voltage issues 
and were disconnected during the experiment, partly justifying the time 
dependent calibration described below. \\

@@

\subsection{Track reconstruction}
~~~~In order to reconstruct tracks we must first select the good hits. The 
first step is the rejection of out-of-time hits and the noise reduction, as 
will be explained in section \ref{sec:noise_rejection}. The second step is the 
spatial reconstruction of the hits, using the extracted drift speed and drift 
paths. For each registered hit, the position of the initial ionization is 
obtained from the recorded time (TDC) and from the position of its pad. In the 
third step, the reconstructed nearby hits are linked together in chains. The 
maximum distance between two adjacent hits must be less than 10.5 mm to chain 
them. Then, the number of hits per chain is required to be greater than 10 hits 
in order to proceed to the fitting step. \\

@@

\section{RTPC calibration}
~~~~Reconstructing a trajectory from the recorded time information of the 
electrons requires a good knowledge of their drift speed and drift paths.  
Also, the gains of the readout pads are required to calculate $\frac{dE}{dx}$ 
from the recorded ADCs. The determination of the drift paths and the gain 
calibration of the RTPC 
require well identified events, so we decided to use elastic scattering 
($e^{4}He \rightarrow e^{4}He$). However, the cross section of 
the elastic process at detectable angles is highly suppressed at higher beam 
energy, so we had specific calibration runs with 1-pass 
beam energies of 1.024 and 1.269 GeV. \\

A GEANT4 simulation for the RTPC has been developed to help reproduce these 
events in order to properly extract the drift paths and the gains. Their 
extractions are based on comparing the experimentally identified $^4He$ elastic 
tracks to the GEANT4 simulated ones. Indeed, the kinematics of each elastic 
$^4He$ can be calculated from the electron measured in CLAS and then simulated 
in GEANT4. The output of the simulation gives an expected trajectory that we then
match to to the measured signals to determine the drift paths. \\

@@

\subsection{Event selection}
~~~~Figure \ref{fig:RTPC_track} shows a drawing of a segment of the RTPC, where a $^4He$ track (in 
green) crosses the volume of the RTPC, producing a chain of ionization points 
within the drift gas. The released electrons follow the drift paths (in black) 
towards the readout board under the effect of the electromagnetic field 
between the anode and the cathode. The electrons released close to the cathode 
take the maximum drift time ($TDC_{max}$) to reach the readout pads, while the 
electrons released close to the anode take the minimum time ($TDC_{min} = 15$ the 
trigger time), also called time offset, which is identical 
for all readout channels. In our convention, the distance between the 
first ionization point in the chain and the cathode is labelled as $sdist$, 
while the distance between the last point and the anode is labelled as 
$edist$.\\

@@

~~~~A good RTPC track passes the following requirements:
\begin{itemize}
\item The number of active pads in the track is greater than 3.

@@

\begin{figure}[tbp]
\begin{minipage}[c]{.46\linewidth}
\hspace{-0.3in}
\includegraphics[height=6.5cm]{fig_rtpc/rtpc_sdist.png}
\caption{$sdist$ distribution for the positive tracks in the RTPC. We set $|sdist|$ < 2.0 mm to select good tracks.} 
\vspace{0.3in}
\label{fig:rrtpc_sdist}
\end{minipage} \hfill
\begin{minipage}[c]{.46\linewidth}
\hspace{-0.3in}
\includegraphics[height=6.5cm]{fig_rtpc/rtpc_edist.png}
\caption{$edist$ distribution for the positive tracks in the RTPC. We require -1.0 mm < $edist$ < 5.0 mm to select the good tracks.}
\vspace{+0.2in}
\label{fig:rrtpc_edist}
\end{minipage}
\end{figure}

@@

\subsubsection{Elastic scattering event selection}
   The elastic process on $^{4}He$ is defined as:
\begin{equation}
\centering
e(k) + ^{4}He(p) \rightarrow e(k') + ^{4}He(p')
\end{equation}
with the symbols in parenthesis representing the four-momenta of the particles. \\

 In addition to the good-track requirements described above, to select elastic events we impose further constraints. The co-planarity between the scattered electron and the recoil $^4He$ is ensured using a $\Delta \phi$ (=$\phi_{e}-\phi_{^{4}He}$) cut as shown in Figure \ref{fig:delta_phi_elastic}. Like for $\Delta z$, the $\Delta \phi$ 2$\sigma$ cut is dependent on $z$ to account for systematic variations. Our final parametrizations can be found in Appendix \ref{app:RTPC_appendix} as well.\\

An additional elastic cut is performed by comparing the measured $^4He$ polar angle to the calculated one, based on the measured electron in CLAS. Indeed, from momentum conservation the $^4He$ polar angle can be calculated as:
\begin{eqnarray}
\theta^{^4He}_{cal} &=&  sin^{-1}\left(\frac{p_{e'}}{p^{^4He}_{cal}} \cdot \sqrt{1-cos^{2}\theta_{e'}}\right) \\
\text{ with~~} p^{^4He}_{cal} &=& \sqrt{(E_{b} + M_{^4He} - p_{e'})^{2} - M^{2}_{^4He}}
\label{equ:theta_helium}
\end{eqnarray}
where $p_{e'}$ and $\theta_{e'}$ are the electron's measured momentum and polar angle, $E_{b}$ is the beam energy and $M_{^4He}$ is the helium mass (3.727 GeV/c$^{2}$). Figure \ref{fig:delta_theta_elastic} shows the $\Delta \theta$ ($\theta^{^4He}_{cal}$ - $\theta^{^4He}_{meas}$) distribution versus the $^4He$ z-vertex. No significant difference was observed between the two modules regarding this quantity. The obtained parametrization of the mean and the width of the $\Delta \theta$  distribution can be found in Appendix \ref{app:RTPC_appendix}.\\

@@

Constructing the invariant mass is one way to verify the efficiency of the previously described selection procedures. For a helium target, the invariant mass (W) is:\\
\begin{equation}
 W = \sqrt{-2E_{b}p_{e^{'}}(1-cos(\theta_{e^{'}})) + M_{^{4}He}^2 + 2(E_{b}-p_{e^{'}}) M_{^{4}He}}.
\end{equation}
with the conventions for the variables as in equation \ref{equ:theta_helium}.  
Figure \ref{fig:W_elastic} shows the W distribution of the events that have a 
good track in the RTPC and the identified elastic events. One can see the good 
agreement between the helium-4 real mass and the mean value of the identified 
elastic events. The difference in width is related to differences in 
resolution in CLAS.
The large effect is due to the reaction products being emitted back to 
back resulting in a small overall acceptance for the process. This result is
therefore very sensitive to the local 
variations of the resolution, which can be rather important in CLAS.\\

@@

\subsubsection{Checking the initial good track requirements}
~~~~In this section, we study some of the good track requirements 
presented in section \ref{good_track_req}. We apply here all the good track 
and elastic requirements but the one related to the ploted variable.

@@

\subsubsection{Drift speed parametrization}
~~~~The electrons follow their drift paths with a certain speed, named drift 
speed. This speed is affected by the experimental conditions, such as the 
variations in the magnetic field or in the gas composition. We can measure this 
speed using the tracks detected in the RTPC, as it is explained in this 
section.\\

In our TPC, the electrons released close to the cathode take the maximum drift 
time ($TDC_{max}$) to reach the readout pads while the geometrical symmetry 
along the RTPC ensures that these electrons always travel the same distance.  
Therefore, by identifying the $TDC_{max}$, the drift speed can be deduced.  
Figure \ref{fig:TDC_profile} shows the time profile of the collected hits for 
the detected good tracks. The experimental drift time ranges between the 
trigger time (TDC$_{min}$ = 15 TDCs) to 75 TDCs. The time profile shows an 
expected dropping edge at high TDCs due to the geometrical constraints. In 
order to avoid the statistical effects in determining the  $TDC_{max}$, we 
define a value named as $TDC_{max/2}$ at which the dropping edge passes to half 
of the maximum number of hits in each hits-time profile. $TDC_{max/2}$ is 
inversely proportional to the drift speed.\\

We measure the $TDC_{max/2}$ along the length of the RTPC to take into account 
the variations in the electric and magnetic fields. The results can be seen in 
figure \ref{fig:RunNumber_61551_TDCmax_Zslice}, where a clear variation of the 
drift speed along the RTPC was observed. \\ \begin{figure}[tbp]
\centering
\vspace{-0.1in}
\includegraphics[scale=0.35]{fig_rtpc/TDC_profile.png}
\caption{Time profile of the collected hits for the RTPC good tracks in one 
   experimental run, run 61511, integrated over the full longitudinal length 
(z) of the RTPC. The $TDC_{max/2}$ is the time at which the dropping edge 
passes half the maximum number of hits.}
\label{fig:TDC_profile}
\end{figure}

We also monitor the $TDC_{max/2}$ evolution during the time of the experiment, 
to take into account variations. The EG6 experiment has recorded data with 
different electron-beam energies: 1.204, 5.7, 6.064 and 1.269 GeV. The 5.7 and 
6.064 data sets have low elastic cross sections compared to the 1.206 and 1.269 
data sets. For this reason, we use all the RTPC good tracks and not only the 
elastic ones for these. Before checking the stability of $TDC_{max/2}$ over the 
experimental running period, we want to check the feasibility of using all good 
tracks for the purpose of parametrizing the drift speed. In this check, the 
$TDC_{max/2}$ identified using the good tracks of the 1.206-GeV data set is 
compared to the one identified from the elastic events of the same data sets.  
The result is shown in figure \ref{fig:TDCmax_ratio}, in which the ratio 
between the two $TDC_{max/2}$ is plotted as a function of the longitudinal 
position along the RTPC. This ratio is consistent with 1, within 1$\%$. Thus we 
can conclude that using the hits of the collected good tracks is a good 
approximation. \\

@@

The variations with time of $TDC_{max/2}$ are shown in figure 
\ref{fig:Drift_run_number}, in which each point is a run of about two hours of 
data taking. One can see a non-negligible variation over the three months of 
data taking due to the changes in the experimental conditions. The main cause 
of these variations is probably the variation of proportions in the drift gas 
mixture which was not perfectly under control due to leaks in the thin parts of 
the detector.\\

As $TDC_{max/2} -15 $ ($= \frac{Drift~path~length}{Drift~speed}$) varies with 
both run number (time) and the geometry of the RTPC ($z$ along the RTPC). We
parametrize the dependence on $z$, hit (pad) position, for each run in the 
form:\\
\begin{equation}
TDC_{max/2} (z)= p_{0} +  p_{1} * e^{p_{2}*(z-p_{3})^{2}}.
\label{equ:drift_speed_fit}
\end{equation}
Then fits are performed for these parameters as functions of the run number, 
such that the drift speed parametrization depends on the run and $z$ along the 
RTPC. The numeric parametrizations of $p_{0}$, $p_{1}$, $p_{2}$, and $p_{3}$ 
can be found in Appendix \ref{app:RTPC_appendix}, table \ref{Table:TDCmax}. \\

These functions are implemented in our reconstruction codes, and, as a result 
of adapting the drift speed with these functions, more good tracks and elastic 
events were identified comparing to the previous calibration set. Figure 
\ref{fig:Drift_run_number} also shows the gain percentages of the good tracks 
(GT) and the elastic events (El) in few runs. One can see a gradual improvement 
with time because of giving the right drift speed, which together with the 
drift paths gives more precise reconstructed hits chain. Therefore, the 
parameters of the reconstructed tracks ($sdist$, $edist$ ... etc) are more likely to pass the 
selection requirements and so more tracks are found to be good. \\

@@

\subsubsection{Drift paths parametrization}
~~~~A drift path is the trajectory that an electron follows after being 
released in the drift region. The standard software to calculate these paths is 
the MAGBOLTZ program \cite{MAGBOLTZ}. This program requires precise knowledge 
of the experimental conditions, such as the detector's geometry, the exact 
composition of the drift gas and the applied electric and magnetic fields. Thus 
any variation in these conditions would give inaccurate calculated drift paths.  
In the EG6 experiment, MAGBOLTZ has been used to calculate a first set of drift 
paths, but it was observed for the drift speed that we do not have a perfect 
knowledge of the RTPC's conditions. Therefore, the first set of the drift paths 
is not accurate. For this reason, we chose to calculate the drift paths in a 
way that does not require an exact knowledge of the conditions in the chamber 
and is instead based on our experimental data.  \\

In our alternative method, we use the identified elastic ($e ~ ^{4}He 
\rightarrow e ~ ^{4}He$) events to extract the drift paths. For these events, 
the kinematics of the recoil $^{4}He$ are calculated from the scattered 
electron. Then the electrons' drift paths can be extracted from comparing the 
experimentally measured hits in the RTPC to the trajectory of a similar 
simulated $^{4}He$. In this technique, the drift paths are obtained 
independently of our knowledge of the exact conditions in the RTPC, in 
particular the electric field and the gas mixture have no impact. \\

Due to the magnetic field, the drift paths are not linear. This is handled by 
the following procedure, in which we perform the drift paths' extraction in 
two passes. In the first pass, we assume a linear correlation between the 
radius of emission $R$ (we work here in radial coordinates: $R$ and $\phi$) and the 
drift time, to link the GEANT4 hits to the measured hits. Through this association, we obtain an 
initial set of drift paths in the form of a $\Delta \phi$ distribution as function of time. We define $\Delta \phi$ the 
difference between the $\phi$ of the simulated hits and the $\phi$ of the pad 
which measured a hit. In the second pass, we refine the initial 
correlation between the radius and the time using the initial drift paths, and 
we extract final drift paths. To take into account the variations of the magnetic 
field and therefore the varying Lorentz angle, the extraction 
is made in $z$ bins along the RTPC. \\

The procedure steps are:\\
\begin{itemize}
\item We first select a sample of well identified elastic events in our experimental data. Using the
   kinematics of the scattered electrons in these events, we generate in GEANT4
   a sample of events identical to the measured ones. 
\item  For each couple of events (the experimental one and its simulated equivalent)
   we make a linear correlation between $R$, from simulation, and TDC from 
   real data, to link the GEANT4 hits to the measured ones. In our TPC, the 
   maximum $R$ equals 60 mm (close to the anode) at the trigger time, 
   (TDC$_{min}$ = 15 TDCs), and the minimum R equals 30 mm (close to the 
   cathode) at maximum TDC ($TDC_{max}$). Thus, the linear correlations can be 
   written as:
\begin{equation}
   R(TDC) = \frac{60-30}{15-TDC_{max}} (TDC-15) + 60.
\label{equ:R_TDC}
\end{equation}
We apply 3 TDCs-wide windows around the linear R-TDC correlation to associate 
the GEANT4 simulated hits to measured hits. Figure \ref{fig:R_correlation} 
shows the distribution of $R$ versus TDC for the selected simulated hits in a 
single $z$ bin.

@@

We performed a third pass in which the R(TDC) correlation is refined using the 
drift paths from the second pass and new drift paths were extracted. As a 
result, we observed no difference between the drift paths of the second and the 
third passes. In other words, extracting the drift paths from two iterations 
gives us stable drift paths. \\

@@

~~~~It is important to note that the drift paths have shown strong sensitivity 
to the other RTPC calibrations. Therefore, we performed iterative extractions 
of the drift paths and, as a results, obtained more good tracks with a signal 
to background ratio largely improved. Figure \ref{fig:sdist_comp_pass1v1_v6} 
shows an illustration of the improvements in terms of the $sdist$ parameter 
using the same initial collected data with an old drift paths and a newer set.

>>>>

At 6 GeV incident electron beam energy, the recoil $^{4}He$ nuclei from 
the coherent DVCS channel, have an average momentum around 200 MeV/c. CLAS 
cannot detect such low energy particles, for exemple protons are detected with 
a threshold of about 250 MeV/c, which will be even higher for helium. In order 
to detect such low-energy recoils a Radial TPC (RTPC), similar to the one used 
during the BoNuS run \cite{BONUS}, was added to CLAS. In 2009, the CLAS-EG6 
group \cite{proposal} has built this RTPC, figure \ref{fig:RTPC2} shows a 
close photo of the RTPC, while figure \ref{fig:RTPC_xscetion} shows a 
schematic drawing of its internal structure. It is a 200~mm-long and 
150~mm-diameter TPC, composed of two electronically separated modules, each 
covering about 150$^{\circ}$ of the azimuthal angle. The RTPC tracks the 
charged particles through the ionization centers they leave in the gas-filled 
detection volume. Then, the ionized electrons drift under the effect of an 
electric field and are amplified by a gas electron multiplier system before to 
be detected on collection pads.

@@

The Ne-DME gas mixture has been chosen as the drift gas because of its 
low-diffusion characteristics and small Lorentz angles (the angles between the 
drift direction of electrons under the influence of magnetic field and the 
direction of the electric field). These characteristics minimize the changes in 
the drift velocity of the ionization electrons \cite{DME}.

The GEMs amplify the ionized electrons to produce measurable signals. Figure 
\ref{fig:GEMs} shows a microscopic photo of a GEM. It is made from an insulator 
(Kapton) sandwiched between two copper layers. The mesh of each GEM layer is 
chemically etched with 50-$\mu$m-diameter holes in double-conical cross section 
shapes, as can be seen from the schematic plot on the right.

@@

The electrons amplification is achieved through the holes via the strong 
electric field, that is generated by a 400-V potential difference between the 
two copper layers. Such a strong field leads to high ionization of the initial 
electrons and therefore amplification of the signals. Furthermore, an 
additional potential difference of 150 V is set between each two successive GEM 
layers to push the amplified electrons towards the readout board. In this 
configuration, the gain of each GEM layer is of the order of 100.

@@

\subsection{Working principle}
When a charged particle traverses a gas, it ionizes the gas-atoms along its 
trajectory. In a TPC, the electrons released in the ionization drift towards 
the readout board under the effect of an applied electric field. The drift 
velocity depends on the gas mixture, and on the electric ($\vec{E}$) and 
magnetic ($\vec{B}$) fields. The recorded time of the electrons provides 
information on how far the initial ionizations happened in the drift region, 
leading to reconstruct the original points of ionization, while the recorded 
ADCs give the deposited energy.

In our TPC the cathode and the anode are two cylinders. Thus, the generated 
$\vec{E}$ field in the drift region has a purely radial components, 
perpendicular to the beam line, with a magnitude around 500 V/cm. However, the 
two-cylinders configuration produces small gradient components at the sides of 
the RTPC. This issue was solved by installing field cages at the ends of the 
RTPC to keep a regular $\vec{E}$ field. These caused some high voltage issues 
and were disconnected during the experiment, partly justifying the time 
dependent calibration described below.

@@

\subsection{Track reconstruction}
In order to reconstruct tracks we must first select the good hits. The 
first step is the rejection of out-of-time hits and the noise reduction, as 
will be explained in section \ref{sec:noise_rejection}. The second step is the 
spatial reconstruction of the hits, using the extracted drift speed and drift 
paths. For each registered hit, the position of the initial ionization is 
obtained from the recorded time (TDC) and from the position of its pad. In the 
third step, the reconstructed nearby hits are linked together in chains. The 
maximum distance between two adjacent hits must be less than 10.5 mm to chain 
them. Then, the number of hits per chain is required to be greater than 10 hits 
in order to proceed to the fitting step.

@@

\section{RTPC calibration}
Reconstructing a trajectory from the recorded time information of the 
electrons requires a good knowledge of their drift speed and drift paths.  
Also, the gains of the readout pads are required to calculate $\frac{dE}{dx}$ 
from the recorded ADCs. The determination of the drift paths and the gain 
calibration of the RTPC 
require well identified events, so we decided to use elastic scattering 
($e^{4}He \rightarrow e^{4}He$). However, the cross section of 
the elastic process at detectable angles is highly suppressed at higher beam 
energy, so we had specific calibration runs with 1-pass 
beam energies of 1.024 and 1.269 GeV.

A GEANT4 simulation for the RTPC has been developed to help reproduce these 
events in order to properly extract the drift paths and the gains. Their 
extractions are based on comparing the experimentally identified $^4He$ elastic 
tracks to the GEANT4 simulated ones. Indeed, the kinematics of each elastic 
$^4He$ can be calculated from the electron measured in CLAS and then simulated 
in GEANT4. The output of the simulation gives an expected trajectory that we then
match to to the measured signals to determine the drift paths.

@@

\subsection{Event selection}
Figure \ref{fig:RTPC_track} shows a drawing of a segment of the RTPC, 
where a $^4He$ track (in 
green) crosses the volume of the RTPC, producing a chain of ionization points 
within the drift gas. The released electrons follow the drift paths (in black) 
towards the readout board under the effect of the electromagnetic field 
between the anode and the cathode. The electrons released close to the cathode 
take the maximum drift time ($TDC_{max}$) to reach the readout pads, while the 
electrons released close to the anode take the minimum time ($TDC_{min} = 15$ the 
trigger time), also called time offset, which is identical 
for all readout channels. In our convention, the distance between the 
first ionization point in the chain and the cathode is labelled as $sdist$, 
while the distance between the last point and the anode is labelled as 
$edist$.

@@

A good RTPC track passes the following requirements:
\begin{itemize}
\item The number of active pads in the track is greater than 3.

@@

\begin{figure}[tbp]
\begin{minipage}[c]{.46\linewidth}
\hspace{-0.3in}
\includegraphics[height=6.5cm]{fig_rtpc/rtpc_sdist.png}
\caption{$sdist$ distribution for the positive tracks in the RTPC. We set $|sdist| < 2.0$ mm to select good tracks.} 
\vspace{0.3in}
\label{fig:rrtpc_sdist}
\end{minipage} \hfill
\begin{minipage}[c]{.46\linewidth}
\hspace{-0.3in}
\includegraphics[height=6.5cm]{fig_rtpc/rtpc_edist.png}
\caption{$edist$ distribution for the positive tracks in the RTPC. We require -1.0 mm < $edist$ < 5.0 mm to select the good tracks.}
\vspace{+0.2in}
\label{fig:rrtpc_edist}
\end{minipage}
\end{figure}

@@

\subsubsection{Elastic scattering event selection}
   The elastic process on $^{4}He$ is defined as:
\begin{equation}
\centering
e(k) + ^{4}He(p) \rightarrow e(k') + ^{4}He(p')
\end{equation}
with the symbols in parenthesis representing the four-momenta of the particles.

 In addition to the good-track requirements described above, to select elastic events we impose further constraints. The co-planarity between the scattered electron and the recoil $^4He$ is ensured using a $\Delta \phi$ (=$\phi_{e}-\phi_{^{4}He}$) cut as shown in Figure \ref{fig:delta_phi_elastic}. Like for $\Delta z$, the $\Delta \phi$ 2$\sigma$ cut is dependent on $z$ to account for systematic variations. Our final parametrizations can be found in Appendix \ref{app:RTPC_appendix} as well.

An additional elastic cut is performed by comparing the measured $^4He$ polar angle to the calculated one, based on the measured electron in CLAS. Indeed, from momentum conservation the $^4He$ polar angle can be calculated as:
\begin{eqnarray}
\theta^{^4He}_{cal} &=&  sin^{-1}\left(\frac{p_{e'}}{p^{^4He}_{cal}} \cdot \sqrt{1-cos^{2}\theta_{e'}}\right) \\
\text{ with~~} p^{^4He}_{cal} &=& \sqrt{(E_{b} + M_{^4He} - p_{e'})^{2} - M^{2}_{^4He}}
\label{equ:theta_helium}
\end{eqnarray}
where $p_{e'}$ and $\theta_{e'}$ are the electron's measured momentum and polar angle, $E_{b}$ is the beam energy and $M_{^4He}$ is the helium mass (3.727 GeV/c$^{2}$). Figure \ref{fig:delta_theta_elastic} shows the $\Delta \theta$ ($\theta^{^4He}_{cal}$ - $\theta^{^4He}_{meas}$) distribution versus the $^4He$ z-vertex. No significant difference was observed between the two modules regarding this quantity. The obtained parametrization of the mean and the width of the $\Delta \theta$  distribution can be found in Appendix \ref{app:RTPC_appendix}.

@@

Constructing the invariant mass is one way to verify the efficiency of the previously described selection procedures. For a helium target, the invariant mass (W) is:
\begin{equation}
 W = \sqrt{-2E_{b}p_{e^{'}}(1-cos(\theta_{e^{'}})) + M_{^{4}He}^2 + 2(E_{b}-p_{e^{'}}) M_{^{4}He}}.
\end{equation}
with the conventions for the variables as in equation \ref{equ:theta_helium}.  
Figure \ref{fig:W_elastic} shows the W distribution of the events that have a 
good track in the RTPC and the identified elastic events. One can see the good 
agreement between the helium-4 real mass and the mean value of the identified 
elastic events. The difference in width is related to differences in 
resolution in CLAS.
The large effect is due to the reaction products being emitted back to 
back resulting in a small overall acceptance for the process. This result is
therefore very sensitive to the local 
variations of the resolution, which can be rather important in CLAS.

@@

\subsubsection{Checking the initial good track requirements}
In this section, we study some of the good track requirements 
presented in section \ref{good_track_req}. We apply here all the good track 
and elastic requirements but the one related to the ploted variable.

@@

\subsubsection{Drift speed parametrization}
The electrons follow their drift paths with a certain speed, named drift 
speed. This speed is affected by the experimental conditions, such as the 
variations in the magnetic field or in the gas composition. We can measure this 
speed using the tracks detected in the RTPC, as it is explained in this 
section.

In our TPC, the electrons released close to the cathode take the maximum drift 
time ($TDC_{max}$) to reach the readout pads while the geometrical symmetry 
along the RTPC ensures that these electrons always travel the same distance.  
Therefore, by identifying the $TDC_{max}$, the drift speed can be deduced.  
Figure \ref{fig:TDC_profile} shows the time profile of the collected hits for 
the detected good tracks. The experimental drift time ranges between the 
trigger time (TDC$_{min}$ = 15 TDCs) to 75 TDCs. The time profile shows an 
expected dropping edge at high TDCs due to the geometrical constraints. In 
order to avoid the statistical effects in determining the  $TDC_{max}$, we 
define a value named as $TDC_{max/2}$ at which the dropping edge passes to half 
of the maximum number of hits in each hits-time profile. $TDC_{max/2}$ is 
inversely proportional to the drift speed.

We measure the $TDC_{max/2}$ along the length of the RTPC to take into account 
the variations in the electric and magnetic fields. The results can be seen in 
figure \ref{fig:RunNumber_61551_TDCmax_Zslice}, where a clear variation of the 
drift speed along the RTPC was observed.

\begin{figure}[tbp]
\centering
\vspace{-0.1in}
\includegraphics[scale=0.35]{fig_rtpc/TDC_profile.png}
\caption{Time profile of the collected hits for the RTPC good tracks in one 
   experimental run, run 61511, integrated over the full longitudinal length 
(z) of the RTPC. The $TDC_{max/2}$ is the time at which the dropping edge 
passes half the maximum number of hits.}
\label{fig:TDC_profile}
\end{figure}

We also monitor the $TDC_{max/2}$ evolution during the time of the experiment, 
to take into account variations. The EG6 experiment has recorded data with 
different electron-beam energies: 1.204, 5.7, 6.064 and 1.269 GeV. The 5.7 and 
6.064 data sets have low elastic cross sections compared to the 1.206 and 1.269 
data sets. For this reason, we use all the RTPC good tracks and not only the 
elastic ones for these. Before checking the stability of $TDC_{max/2}$ over the 
experimental running period, we want to check the feasibility of using all good 
tracks for the purpose of parametrizing the drift speed. In this check, the 
$TDC_{max/2}$ identified using the good tracks of the 1.206-GeV data set is 
compared to the one identified from the elastic events of the same data sets.  
The result is shown in figure \ref{fig:TDCmax_ratio}, in which the ratio 
between the two $TDC_{max/2}$ is plotted as a function of the longitudinal 
position along the RTPC. This ratio is consistent with 1, within 1$\%$. Thus we 
can conclude that using the hits of the collected good tracks is a good 
approximation.

@@

The variations with time of $TDC_{max/2}$ are shown in figure 
\ref{fig:Drift_run_number}, in which each point is a run of about two hours of 
data taking. One can see a non-negligible variation over the three months of 
data taking due to the changes in the experimental conditions. The main cause 
of these variations is probably the variation of proportions in the drift gas 
mixture which was not perfectly under control due to leaks in the thin parts of 
the detector.

As $TDC_{max/2} -15 $ ($= \frac{Drift~path~length}{Drift~speed}$) varies with 
both run number (time) and the geometry of the RTPC ($z$ along the RTPC). We
parametrize the dependence on $z$, hit (pad) position, for each run in the 
form:
\begin{equation}
TDC_{max/2} (z)= p_{0} +  p_{1} * e^{p_{2}*(z-p_{3})^{2}}.
\label{equ:drift_speed_fit}
\end{equation}
Then fits are performed for these parameters as functions of the run number, 
such that the drift speed parametrization depends on the run and $z$ along the 
RTPC. The numeric parametrizations of $p_{0}$, $p_{1}$, $p_{2}$, and $p_{3}$ 
can be found in Appendix \ref{app:RTPC_appendix}, table \ref{Table:TDCmax}.

These functions are implemented in our reconstruction codes, and, as a result 
of adapting the drift speed with these functions, more good tracks and elastic 
events were identified comparing to the previous calibration set. Figure 
\ref{fig:Drift_run_number} also shows the gain percentages of the good tracks 
(GT) and the elastic events (El) in few runs. One can see a gradual improvement 
with time because of giving the right drift speed, which together with the 
drift paths gives more precise reconstructed hits chain. Therefore, the 
parameters of the reconstructed tracks ($sdist$, $edist$ ... etc) are more likely to pass the 
selection requirements and so more tracks are found to be good.

@@

\subsubsection{Drift paths parametrization}
A drift path is the trajectory that an electron follows after being 
released in the drift region. The standard software to calculate these paths is 
the MAGBOLTZ program \cite{MAGBOLTZ}. This program requires precise knowledge 
of the experimental conditions, such as the detector's geometry, the exact 
composition of the drift gas and the applied electric and magnetic fields. Thus 
any variation in these conditions would give inaccurate calculated drift paths.  
In the EG6 experiment, MAGBOLTZ has been used to calculate a first set of drift 
paths, but it was observed for the drift speed that we do not have a perfect 
knowledge of the RTPC's conditions. Therefore, the first set of the drift paths 
is not accurate. For this reason, we chose to calculate the drift paths in a 
way that does not require an exact knowledge of the conditions in the chamber 
and is instead based on our experimental data.

In our alternative method, we use the identified elastic ($e ~ ^{4}He 
\rightarrow e ~ ^{4}He$) events to extract the drift paths. For these events, 
the kinematics of the recoil $^{4}He$ are calculated from the scattered 
electron. Then the electrons' drift paths can be extracted from comparing the 
experimentally measured hits in the RTPC to the trajectory of a similar 
simulated $^{4}He$. In this technique, the drift paths are obtained 
independently of our knowledge of the exact conditions in the RTPC, in 
particular the electric field and the gas mixture have no impact.

Due to the magnetic field, the drift paths are not linear. This is handled by 
the following procedure, in which we perform the drift paths' extraction in 
two passes. In the first pass, we assume a linear correlation between the 
radius of emission $R$ (we work here in radial coordinates: $R$ and $\phi$) and the 
drift time, to link the GEANT4 hits to the measured hits. Through this association, we obtain an 
initial set of drift paths in the form of a $\Delta \phi$ distribution as function of time. We define $\Delta \phi$ the 
difference between the $\phi$ of the simulated hits and the $\phi$ of the pad 
which measured a hit. In the second pass, we refine the initial 
correlation between the radius and the time using the initial drift paths, and 
we extract final drift paths. To take into account the variations of the magnetic 
field and therefore the varying Lorentz angle, the extraction 
is made in $z$ bins along the RTPC.

The procedure steps are:
\begin{itemize}
\item We first select a sample of well identified elastic events in our experimental data. Using the
   kinematics of the scattered electrons in these events, we generate in GEANT4
   a sample of events identical to the measured ones. 
\item  For each couple of events (the experimental one and its simulated equivalent)
   we make a linear correlation between $R$, from simulation, and TDC from 
   real data, to link the GEANT4 hits to the measured ones. In our TPC, the 
   maximum $R$ equals 60 mm (close to the anode) at the trigger time, 
   (TDC$_{min}$ = 15 TDCs), and the minimum R equals 30 mm (close to the 
   cathode) at maximum TDC ($TDC_{max}$). Thus, the linear correlations can be 
   written as:
\begin{equation}
   R(TDC) = \frac{60-30}{15-TDC_{max}} (TDC-15) + 60.
\label{equ:R_TDC}
\end{equation}
We apply 3 TDCs-wide windows around the linear R-TDC correlation to associate 
the GEANT4 simulated hits to measured hits. Figure \ref{fig:R_correlation} 
shows the distribution of $R$ versus TDC for the selected simulated hits in a 
single $z$ bin.

@@

We performed a third pass in which the R(TDC) correlation is refined using the 
drift paths from the second pass and new drift paths were extracted. As a 
result, we observed no difference between the drift paths of the second and the 
third passes. In other words, extracting the drift paths from two iterations 
gives us stable drift paths.

@@

It is important to note that the drift paths have shown strong sensitivity 
to the other RTPC calibrations. Therefore, we performed iterative extractions 
of the drift paths and, as a results, obtained more good tracks with a signal 
to background ratio largely improved. Figure \ref{fig:sdist_comp_pass1v1_v6} 
shows an illustration of the improvements in terms of the $sdist$ parameter 
using the same initial collected data with an old drift paths and a newer set.
